quollr: An R Package for Visualizing $2-D$ Models from Non-linear Dimension Reductions in High Dimensional Space

### 2024-06-22 paper-quollr.Rmd
Add details regarding the binning data

The number of bins along the x-axis, the ratio of the ranges of the original embedding components, the buffer amount as a proportion of data, and if `is_rm_lwd_hex = TRUE`, benchmark value to remove low density hexagons are parameters that will be determined within `fit_highd_model`, if they are not provided. They are created as they are needed throughout the following example.

@@

Constructing the $2\text{-}D$ model mainly contains (i) scaling, (ii) computing hexagon grid configurations, (iii) binning, and (iv) indicating neighbors by line segments connecting centroids.

@@

```{r, echo=TRUE}
calc_bins_y(
  bin1 = 10, 
  r2 = r2
  )
```

@@

```{r, echo=TRUE}
hex_binning(
  data = s_curve_noise_umap_scaled, 
  bin1 = 10, 
  r2 = r2
  )
```

@@

## Model parameters

<!--discuss about default settings-->

### Choice of bins

### Removal of low-density bins

```{r, echo=TRUE, eval=FALSE}

find_low_dens_hex(
  df_bin_centroids_all = df_bin_centroids, 
  bin1 = num_bins_x, 
  df_bin_centroids_low = df_bin_centroids_low
  )

### Removing long edges

@@

This paper presents the R package `quollr` to develop a way to take the fitted model, as represented by the positions of points in 2D, and turn it into a high-dimensional wireframe to overlay on the data, viewing it with a tour.

@@

This new tool provides an effective start point for automatically creating regular hexagons and help to evaluate which NLDR technique and which hyperparameter choice gives the most accurate representation of $p-D$ data.

>>>>

The number of bins along the x-axis, the ratio of the ranges of the original embedding components, the buffer amount as a proportion of data, and if `is_rm_lwd_hex = TRUE`, benchmark value to remove low density hexagons are parameters that will be determined within `fit_highd_model()`, if they are not provided. They are created as they are needed throughout the following example. The function `fit_highd_model()` provides the fitted model in $2\text{-}D$, and $p\text{-}D$.

@@

Constructing the $2\text{-}D$ model mainly contains (i) binning data, (ii) obtaining bin centroids, and (iii) indicating neighbors by line segments connecting centroids.

@@

The configurations of the hexagonal grid is defined by the number of bins in each direction. To find the number of bins along the y-axis, `calc_bins_y()` is used. This function takes as input the number of bins along the x-axis, the ratio of the ranges of the original embedding components, and the buffer amount as a proportion of the data. Additionally, this function provides the bin width.

```{r, echo=TRUE}
calc_bins_y(
  bin1 = 12, 
  r2 = r2,
  q = 0.07
  )
```

@@

Points are allocated to the bins they fall into based on the nearest centroid. The main steps of the hexagonal binning algorithm can be executed using the `hex_binning()` function, or they can be run separately for greater flexibility. The parameters used within `hex_binning()` include the scaled NLDR data, the number of bins along the x-axis, the ratio of the ranges of the original embedding components, and the buffer amount as a proportion of the data. The output is an object of the `hex_bin_obj` class, which contains the number of bins in each direction, the coordinates of the hexagonal grid starting point, the details of bin centroids, the coordinates of bins, embedding components with their corresponding hexagon IDs, hex bins with their corresponding standardized counts, the total number of bins, the number of non-empty bins, and the points within each hexagon.

```{r, echo=TRUE}
hex_binning(
  data = s_curve_noise_umap_scaled, 
  bin1 = 12, 
  r2 = r2,
  q = 0.07
  )
```

### Remove low-density hexagons

@@

@@

This paper presents the R package `quollr` to develop a way to take the fitted model, as represented by the positions of points in $2\text{-}D$, and turn it into a high-dimensional wireframe to overlay on the data, viewing it with a tour.

@@

This new tool provides an effective start point for automatically creating regular hexagons and help to evaluate which NLDR technique and which hyperparameter choice gives the most accurate representation of $p\text{-}D$ data.
